# List of experiments
Define lists that contain the plot settings for different experiments


### B11 vs B12 lysis buffers
Compare buffer B11 and B12.
Use batches MSEB0006 (caecum) and MSEB0010 (colon), from the focal (adult) chicken. 
For the colon, use the samples that were amplified with 15 PCR cycles instead of 19 (due to the latter's low quality).
```{r data_set_buffers, warning=FALSE, comments="", message=FALSE}
plot_params_buffers <- list(
  filter_conditions = list(
    quote(section != "Ileum"),
    quote(cycles < 16),
    quote(batch == "MSEB0006" | batch == "MSEB0010")
  ),
  labels_title = "Lysis Buffer",
  facet_formula = ". ~ section + type + buffer", #"batch + section + type ~ ."
  scale_fill_manual_val = c('#ffdf9e','#ffc273'),
  fill_var = "buffer",
  plot_title = "Lysis Buffer trial"
)
```

#### 15 vs 19 PCR cycles
Use the colon samples (MSEB0010). Maybe separate by buffer??
```{r data_set_cycles, message=FALSE, warning=FALSE, comments=""}
plot_params_cycles <- list(
  filter_conditions = list(
    quote(batch == "MSEB0010")
  ),
  labels_title = "PCR cycles",
  facet_formula = "section + type + cycles ~ .", # "batch + section + type ~ ."
  scale_fill_manual_val = c('#ffc273','#e56969'),
  fill_var = "factor(cycles)",
  plot_title = "PCR cycles trial"
)
```

#### Limit of detection trial: Different LMD sizes
Use batch MSEB0014 (caecum). 
```{r data_set_LOD, warning=FALSE, comments="", message=FALSE}
plot_params_LOD <- list(
  filter_conditions = list(
    quote(batch == "MSEB0014")
  ),
  labels_title = "LMD size",
  facet_formula = "type + size ~ .", #"batch + section + type + cryosection ~ ."
  scale_fill_manual_val = c('#ffdf9e','#ffc273','#e56969','#c1558b','#8a49a1','#4f5bd5'),
  fill_var = "factor(size, levels = c(500, 1500, 2500, 5000, 25000, 50000))",
  plot_title = "Limit of detection (LMD size)"
)
```

#### Automation trial
Compare library prep with DreamPrep (MSEB0015) vs manual (MSEB0011) for ceacum of focal chicken
```{r data_set_automation, warning=FALSE, comments="", message=FALSE}
plot_params_automation <- list(
  filter_conditions = list(
    quote(batch == "MSEB0011"|batch == "MSEB0015"), 
    quote(animal == 'G121e')
  ),
  labels_title = "Automation",
  facet_formula = "batch + type + cryosection ~ .", #"batch + section + type + cryosection ~ ."
  scale_fill_manual_val = c('#e56969','#c1558b'),
  fill_var = "batch",
  plot_title = "Automation test"
)
```

#### Full vs. half reaction (library prep with UltraLowV2 Tecan kit)
Compare library prep with full reaction (MSEB0006, MSEB0009, MSEB0010) vs half reaction (MSEB0011, MSEB0012) of focal chicken, ceacum and colon (only low PCR cycles). NB! both buffers.
```{r data_set_protocol, warning=FALSE, comments="", message=FALSE}
plot_params_protocol <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch != "MSEB0014"& batch != "MSEB0015"),
    quote(animal == 'G121e'),
    quote(cycles<16)
  ),
  labels_title = "Protocol",
  facet_formula = "type + section + protocol ~ .", #"type + section + batch ~ ."
  scale_fill_manual_val = c('#c1558b','#8a49a1'),
  fill_var = "protocol",
  plot_title = "Full vs. half reactions"
)
```

#### Ceacum vs colon
Compare colon vs caecum samples of the focal chicken (and only low PCR cycles)
```{r data_set_section, warning=FALSE, comments="", message=FALSE}
plot_params_section <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch == "MSEB0009"|batch == "MSEB0010"|batch == "MSEB0011"|batch == "MSEB0012"),
    quote(animal == 'G121e'),
    quote(cycles<16)
  ),
  labels_title = "Section",
  facet_formula = "type + section ~ .", #"type+ batch ~ ."
  scale_fill_manual_val = c('#8a49a1','#4f5bd5'),
  fill_var = "section",
  plot_title = "Caecum vs colon"
)
```

#### Adult vs young chicken
Compare samples from the focal (adult) chicken vs the younger chicken, for both colon (MSEB0012) and caecum (MSEB0011).
```{r data_set_animal, warning=FALSE, comments="", message=FALSE}
plot_params_animal <- list(
  filter_conditions = list(
    quote(batch == "MSEB0011"|batch == "MSEB0012") 
  ),
  labels_title = "Animal",
  facet_formula = "type + section + animal ~ .", #"type+ batch + section + animal ~ ."
  scale_fill_manual_val = c('#ffc273','#c1558b'),
  fill_var = "animal",
  plot_title = "Adult vs young chicken"
)
```

#### LMD collection attemps
Compare samples coloured by the number of attempts to collect the LMD sample. LOD trial excluded.
```{r data_set_collection_attempts, warning=FALSE, comments="", message=FALSE}
plot_params_collection_attempts <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch != "MSEB0014"),
    quote(animal == 'G121e'),
    quote(cycles<16),
    quote(collection_attempts>0)
  ),
  labels_title = "Collection attempts",
  facet_formula = "type + section + collection_attempts ~ .", #"type + section + batch ~ ."
  scale_fill_manual_val = c('#ffdf9e','#ffc273','#e56969','#c1558b','#8a49a1','#4f5bd5'),
  fill_var = "factor(collection_attempts)",
  plot_title = "LMD collection attempts"
)
```

#### LMD collection success
Compare samples coloured by the LMD success jugded upon visual inspection of the collection lids. LOD trial excluded.
```{r data_set_collection_success, warning=FALSE, comments="", message=FALSE}
plot_params_collection_success <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch != "MSEB0014"),
    quote(animal == 'G121e'),
    quote(cycles<16),
    quote(collection_attempts>0)
  ),
  labels_title = "Collection_success",
  facet_formula = "type + section + collection ~ .", #"type + section + batch ~ ."
  scale_fill_manual_val = c('#ffc273','#e56969','#c1558b','#8a49a1','#4f5bd5'),
  fill_var = "collection",
  plot_title = "LMD collection success"
)
```

#### LMD collection container
Compare samples coloured by the container they were collected in during the microsection.
```{r data_set_container_collection, warning=FALSE, comments="", message=FALSE}
plot_params_container_collection <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch == "MSEB0009"),
    quote(cycles<16), 
  quote(protocol == "ULV2_Full")
  ),
  labels_title = "Collection_container",
  facet_formula = "type + collection_method ~ .",
  scale_fill_manual_val = c('#c1558b','#4f5bd5'),
  fill_var = "collection_method",
  plot_title = "LMD collection container"
)
```


#### Sections of Ceacum vs colon
Compare colon vs caecum samples of the focal chicken (and only low PCR cycles)
```{r data_set_slices, warning=FALSE, comments="", message=FALSE}
plot_params_slices <- list(
  filter_conditions = list(
    quote(section != "Ileum"), 
    quote(batch == "MSEB0011"|batch == "MSEB0012"),
    quote(animal == 'G121e'),
    quote(cycles<16)
  ),
  labels_title = "Section",
  facet_formula = "type + section + cryosection ~ .", #"type+ batch ~ ."
  scale_fill_manual_val = c('#e56969','#8a49a1'),
  fill_var = "section",
  plot_title = "Caecum vs colon"
)
```

### Define a list of all the experiments/trials settings that you want to plot.

```{r experiment_settings_list, warning=FALSE, comments="", message=FALSE}
plot_params_list <- list(
  plot_buffers = plot_params_buffers, 
  plot_cycles = plot_params_cycles,
  plot_LOD = plot_params_LOD,
  plot_automation = plot_params_automation,
  plot_protocol = plot_params_protocol,
  plot_section = plot_params_section,
  plot_animal = plot_params_animal,
  plot_collection_attempts = plot_params_collection_attempts,
  plot_collection_success = plot_params_collection_success,
  plot_container_collection = plot_params_container_collection,
  plot_slices = plot_params_slices
)

```

  
### Define lists that contain the settings for plotting each statistic.

#### Sequencing depth
This is the total number of sequenced reads.
```{r stat_params_total_sequences, warning=FALSE, comments="", message=FALSE}
stat_params_total_sequences <- list(
  x_var = "total_sequences",
  x_label = "Number of reads",
  x_vline = 10000000,
  stacked = FALSE
)
```

#### Number of sequences after trimming
This is the total number of sequenced reads after trimming the adaptors and low quality sequences.
```{r stat_params_trimmed_sequences, warning=FALSE, comments="", message=FALSE}
stat_params_trimmed_sequences <- list(
  x_var = "total_trimmed_sequences",
  x_label = "Number of trimmed reads",
  x_vline = NULL,
  stacked = FALSE
)
```

#### Number of trimmed sequences
This is the difference between total reads and reads after trimming.
```{r stats_plot_param_total_trimmed, warning=FALSE, comments="", message=FALSE}
prepare_stacked_data <- function(data) {
  data %>%
    mutate(trimmed_reads = total_sequences - total_trimmed_sequences) %>%
    pivot_longer(cols = c(total_trimmed_sequences, trimmed_reads),
                 names_to = "read_type", values_to = "reads") %>%
    mutate(read_type = factor(read_type, levels = c("trimmed_reads", "total_trimmed_sequences")))
}

stat_params_compare_sequences <- list(
  x_var = "total_trimmed_sequences",
  x_label = "Number of trimmed reads",
  x_vline = NULL,
  stacked = TRUE
)
```

#### Percentage (%) of unique sequences
```{r stat_params_unique, warning=FALSE, comments="", message=FALSE}
stat_params_unique <- list(
  x_var = "percent_unique",
  x_label = "% of unique sequences",
  x_vline = 35,
  stacked = FALSE
)
```

#### Percentage (%) of GC content
```{r stat_params_gc, warning=FALSE, comments="", message=FALSE}
stat_params_gc <- list(
  x_var = "percent_gc",
  x_label = "% of GC content",
  x_vline = 60,
  stacked = FALSE
)
```

#### Percentage (%) of host reads
```{r stat_params_host_reads, warning=FALSE, comments="", message=FALSE}
stat_params_host_reads <- list(
  x_var = "reads_mapped_host_percent",
  x_label = "% of host reads",
  x_vline = NULL,
  stacked = FALSE
)
```

#### Percentage (%) of human reads
```{r stat_params_human_reads, warning=FALSE, comments="", message=FALSE}
stat_params_human_reads <- list(
  x_var = "reads_mapped_human_percent",
  x_label = "% of human reads",
  x_vline = 5,
  stacked = FALSE
)
```

#### Percentage (%) of bacterial reads
NB! In the next iteration, it is better to do this by using the counts dataset instead of the statistics file.
```{r stat_params_bacteria_reads, warning=FALSE, comments="", message=FALSE}
stat_params_bacteria_reads <- list(
  x_var = "reads_mapped_bacteria_percent",
  x_label = "% of bacteria reads",
  x_vline = 75,
  stacked = FALSE
)
```

#### % of unmapped reads
```{r stat_params_other_reads, warning=FALSE, comments="", message=FALSE}
stat_params_other_reads <- list(
  x_var = "reads_mapped_other_percent",
  x_label = "% of other reads",
  x_vline = 25,
  stacked = FALSE
)
```

#### Quality score
```{r stat_quality_score, warning=FALSE, comments="", message=FALSE}
stat_quality_score <- list(
  x_var = "quality",
  x_label = "Quality score",
  x_vline = 5,
  stacked = FALSE
)
```


### Define a list of all the statistics settings that you want to plot.
```{r stats_settings_list, warning=FALSE, comments="", message=FALSE}
stat_params_list <- list(
  stat_total_sequences = stat_params_total_sequences,
  stat_trimmed_sequences = stat_params_trimmed_sequences,
  stat_compare_sequences = stat_params_compare_sequences,
  stat_unique = stat_params_unique,
  stat_gc = stat_params_gc,
  stat_host = stat_params_host_reads,
  stat_human = stat_params_human_reads,
  stat_bacteria = stat_params_bacteria_reads,
  stat_other = stat_params_other_reads,
  stat_quality = stat_quality_score
)
```

```{r wrap_params_object, eval=FALSE}
save(plot_params_list,
  stat_params_list,
  file = "data/params.Rdata"
)
```