---
title: "3D'omics | Avian microbiome - Methods MSSM manuscript"
author:
- Antton Alberdi^[University of Copenhagen, antton.alberdi@sund.ku.dk]
- Amalia Bogri^[University of Copenhagen, amalia.bogri@sund.ku.dk]
- Carlotta Pietroni^[University of Copenhagen, carlotta.pietroni@sund.ku.dk]
- Jorge Langa^[University of Copenhagen, jorge.langa@sund.ku.dk]
date: "`r Sys.Date()`"
subtitle: Work in progress
site: bookdown::bookdown_site
documentclass: book
bibliography:
- book.bib
- packages.bib
url: "https://3d-omics.github.io/XXXX"
description: |
  Data analysis code for 3D'omics microdissection methods manuscript (MSSM)
link-citations: true
github-repo: "3d-omics/XXX"
---

```{r knitr_opts, echo=FALSE}
knitr::opts_chunk$set(
  class.source = "script-source",
  class.output = "script-output",
  comment = NA
)
```

# Introduction

This webbook contains all the code used for the analyses of the MSSM methods manuscript.

## Prepare the R environment

### Environment

To reproduce all the analyses locally, clone this repository in your computer using:

```
RStudio > New Project > Version Control > Git
```

And indicating the following git repository:

> https://github.com/3d-omics/ANMI014_chicken_LMD_methods.git

Once the R project has been created, follow the instructions and code chunks shown in this webbook.

### Libraries

The following R packages are required for the data analysis.

```{r load_libraries, warning=FALSE, comments="", message=FALSE}
# Base
library(R.utils)
library(knitr)
library(devtools)
library(tinytable)
library(broom)
library(broom.mixed)
library(rairtable)
library(jsonlite)

# For tree handling
library(ape)
library(phyloseq)
library(phytools)

# For plotting
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(ggnewscale)
library(gridExtra)
library(ggtreeExtra)
library(ggtree)
library(ggh4x)
library(jpeg)
library(patchwork)

# For statistics
library(spaa)
library(vegan)
library(Rtsne)
library(geiger)
library(hilldiv2)
library(distillR)
library(ANCOMBC)
library(lme4)
library(Hmsc)
library(psych)

# For compositional analyses
library(zCompositions)
library(compositions)

# Tidyverse libraries (load dplyr last to avoid conflicts)
library(tidyverse)
library(dplyr) # Explicitly load dplyr last
```

## Notes:
### Installation of modules
Most of the modules can be downloaded from CRAN with 'install.packages('XXX', dependencies = TRUE)'.
Four modules are from Bioconductor. 
Install with: 'install.packages('BiocManager', dependencies = TRUE)
BiocManager::install(c("phyloseq", "ggtreeExtra","ggtree","ANCOMBC"))'
Two modules are from Antton's github. 
Install with: 'install.packages('remotes', dependencies = TRUE)
remotes::install_github('anttonalberdi/hilldiv2')
remotes::install_github('anttonalberdi/distillR')'

### Airtable access:
Currently the metadata is loaded from the 3D'omics airtable. In order to access it you need your own API key (from Antton). set_airtable_api_key('XXXXX', install = TRUE)
